---
title: "R Notebook"
output: html_notebook
---

```{r}
#install.packages("readxl")
library(readxl)
```

DRIVERS FILE

```{r}
#load datatset
drivers <- read_excel("drivers.xlsx", sheet = "Sheet1")

print(head(drivers))
print(nrow(drivers))

#NAs
sum(is.na(drivers))

```

Dataset is loaded properly, no NAs and number of obserations as the excel file.

Now the data will be properly extract:

```{r}
#Online time of each driver
drivers$online_time <- as.numeric(difftime(drivers$offline_datetime,
                                           drivers$arrival_datetime, units="hours"))
#Inter-arrival time for driver (nrow-1 observation!)
drivers$inter_arr <- c(NA,diff(drivers$arrival_datetime)/3600) #in hour
#Arrival coordinates:

#x-axis
drivers$x_coordinate <- as.numeric(gsub("[()]", "", 
                            sapply(strsplit(drivers$initial_location, ", "), "[", 1)))

#y-axis
drivers$y_coordinate <- as.numeric(gsub("[()]", "", 
                            sapply(strsplit(drivers$initial_location, ", "), "[", 2)))

#verify
head(drivers[, c("id", "x_coordinate", "y_coordinate")])


```

ONLINE TIME

```{r}
#histogram
hist(drivers$online_time,
     col = "lightblue",
     border = "black",
     main = "Distribution of Online Time Drivers",
     xlab = "Time online (h)",
     ylab = "Frequency",
     xlim=c(4,9))


#Add red line (paper says uniform between 5 and 8)
abline(v = 5, col = "red", lty = 2, lwd = 2)  
abline(v = 8, col = "red", lty = 2, lwd = 2)

legend("topright", legend = c("Observed Data", "Uniform Limit (5,8)"),
       col = c("lightblue", "red"), lty = c(NA, 2), pch = c(15, NA), lwd = c(NA, 2),
       bg = "white",cex=0.8)
```

The distribution BoxCar proposed is Uniform(5,8). Clearly here we do not observe any data between 5 and 6, other than that the distribution could be a uniform. We will perform test with H0 : Uniform(6,8).

```{r}
#test for online time of drivers

#H0: Uniform(6,8)
a <- 6
b <- 8

#we can perform a CHI-SQUARED TEST, it's possible to assume that each driver online time is independent from other drivers, then we can perform a Chi-Squared test

#We will work with bins with constat prova inside each

#thumb-rule, choos k in order to have expected number of observation in each bin >=5
n_drivers <- nrow(drivers)

#383 observation, seems reasonable to have k = 20

alpha <- 0.10

k <- 20

p <- 1/k

expected_i <- n_drivers * p #greater than 5, thumb rule ok

bins <- seq(a, b, length.out = k + 1)

counting <- cut(drivers$online_time, breaks = bins, include.lowest = TRUE)

O_i <- table(counting)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k-1, lower.tail = TRUE)

chi_stat > chi_quant

print(chi_stat)
print(chi_quant)

#FAIL TO REJECT H0 !!!!!!! ONLINE TIME DRIVERS IS UNIFORM(6,8)
```

INTERARRIVAL TIME FOR DRIVERS

```{r}
hist(drivers$inter_arr,
     col = "lightblue",
     border = "black",
     probability = TRUE,
     breaks=20,
     main = "Distribution of Inter Arrival Times for Drivers",
     xlab = "Inter-arrival time (h)",
     ylab = "Density")

x_vals <- seq(0, max(na.omit(drivers$inter_arr)), length.out = 100)
lines(x_vals, dexp(x_vals, rate = 3), col = "darkblue", lwd = 2)

legend("topright", legend = c("Observed Inter arrival times", 
                              "Exponential Distribution (rate 3)"),
       col = c("lightblue", "darkblue"), lty = c(NA, 1), 
       pch = c(15, NA), lwd = c(NA, 2),
       bg = "white")
```

Blue line is the exponential(3) distribution. Seems a reasonable test.

```{r}
#Again inter-arrival time of drivers can be assumed as independent, so we perform a chi-squared test

lambda <- 3
k <- 20 #as before

p <- 1/k

expected_i <- n_drivers * p

#build intervals with constant prob inside

num_int <- 0:(k-1)

bins<- c(-log( 1 - num_int / k) / lambda, Inf) #build k intervals with const
                                                #prob

#count O_i

counting <- cut(na.omit(drivers$inter_arr), breaks = bins, 
                include.lowest = TRUE)

O_i <- table(counting)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k-1, lower.tail = TRUE)

chi_stat > chi_quant

print(chi_stat)
print(chi_quant)
#REFUSE EXPONENTIAL WITH RATE =3 REJECT H0!!!!

```

```{r}
#Since the distribution of inter arrival times seems to follow an exponential distribution 

lambda_hat <- 1/mean(na.omit(drivers$inter_arr))
cat("lambda MLE is:", lambda_hat,"\n")

#H0 is inter-arrival time is exp(rate=lamda.hat)
k <- 20 #as before

p <- 1/k

expected_i <- n_drivers * p

#build intervals with constant prob inside

num_int <- 0:(k-1)

bins<- c(-log( 1 - num_int / k) / lambda_hat, Inf) #build k intervals with const
                                                #prob

#count O_i

counting <- cut(na.omit(drivers$inter_arr), breaks = bins, 
                include.lowest = TRUE)

O_i <- table(counting)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k-1-1, lower.tail = TRUE) #fix the dof bcs we estimated lambda hat

chi_stat > chi_quant

print(chi_stat)
print(chi_quant)

#SO INTER-ARRIVAL LOCATION FOLLOWS A EXP(rate= lambda_hat (4....))|!H0
```

```{r}
hist(drivers$inter_arr,
     col = "lightblue",
     border = "black",
     probability = TRUE,
     breaks=20,
     main = "Distribution of Inter Arrival Times for Drivers",
     xlab = "Inter-arrival time (h)",
     ylab = "Density")

x_vals <- seq(0, max(na.omit(drivers$inter_arr)), length.out = 100)
lines(x_vals, dexp(x_vals, rate = lambda_hat), col = "darkblue", lwd = 2)
lines(x_vals, dexp(x_vals, rate = 3), col = "red", lwd = 2)

legend("topright", legend = c("Observed Inter arrival times", 
                              paste0("Exponential Distribution (rate ",
                              round(lambda_hat, 2),")"), 
                              "Exponential Distribution (rate 3)"),
       col = c("lightblue", "darkblue","red"), lty = c(NA, 1, 1), 
       pch = c(15, NA,NA), lwd = c(NA, 2,2),
       bg = "white")
```

ARRIVAL LOCATION FOR DRIVERS

```{r}
plot(drivers$x_coordinate, drivers$y_coordinate,
     col = rgb(0, 0, 1, 0.5), pch = 16,
     xlab = "X", ylab = "Y",
     main = "Initial Distribution of Drivers")
grid()

```

Seems more concentrated in the center of the area (center part of Squareshire).

```{r}
#let's test for Uniform( 0, 20 ) * Uniform(0 , 20), assuming independence between location for each driver. We perform chisquared

k <- 5
a <- 0
b <- 20

#bins on x and y
x_seq <- seq(a, b , length.out = k + 1)
y_seq <- seq(a, b , length.out = k + 1)


#count on x and y
x_count <- cut(drivers$x_coordinate, breaks = x_seq, include.lowest = TRUE)
y_count <- cut(drivers$y_coordinate, breaks = y_seq, include.lowest = TRUE)

obs_table <- table(x_count, y_count)

obs_table

expected_i <- sum(obs_table) / (k * k)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k*k-1, lower.tail = TRUE)

chi_stat > chi_quant

print(chi_stat)

print(chi_quant) 

#NOT UNIFORM AT ALL!!!!!
```

Let's see component X and Y separately:

```{r}
#X-axis
hist(drivers$x_coordinate, col = "lightblue", border = "black",
     main = "Distribution of Arrival Location X-axis Drivers", 
     xlab = "X-Location",
     breaks=20,
     ylab = "Frequency")
```

```{r}
library(truncnorm)
library(stats4)
nll <- function(mu, sigma) {
  
  -sum(log(dtruncnorm(drivers$x_coordinate, a = 0, b = 20, mean = mu, sd = sigma)))
}

fit_trunc <- mle(nll, start = list(mu = mean(drivers$x_coordinate), sigma = sd(drivers$x_coordinate)))
summary(fit_trunc)

mu_est <- fit_trunc@coef["mu"]       
sigma_est <- fit_trunc@coef["sigma"]


F_trunc <- function(x, mu, sigma) {
  lower <- 0
  upper <- 20
  (pnorm(x, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma)) /
    (pnorm(upper, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma))
}

ks_result <- ks.test(drivers$x_coordinate, function(x) F_trunc(x, mu_est, sigma_est))
print(ks_result)

# Plot histogram with density overlay in English
hist(drivers$x_coordinate, 
     col = "lightblue", border = "black",
     main = "Initial Driver Distribution on X-axis", 
     xlab = "X-Coordinate", 
     breaks = 20,
     ylab = "Density", 
     probability = TRUE)

# Plot the fitted truncated normal density curve
curve(dtruncnorm(x, a = 0, b = 20, mean = mu_est, sd = sigma_est), 
      from = 0, to = 20, 
      add = TRUE, col = "red", lwd = 2)

# Add legend with a blue square for observed data and red line with parameter estimates
legend("topright", 
       legend = c("Observed Driver Location ", sprintf("Truncated Normal (mu = %.2f, sigma = %.2f)", mu_est, sigma_est)),
       col = c("lightblue", "red"),
       pch = c(15, NA),  
       lty = c(NA, 1),    
       lwd = c(NA, 2),
       cex=0.6)
```

```{r}


```

```{r}
#Y-axis
hist(drivers$y_coordinate, col = "lightblue", border = "black",
     main = "Distribution of Arrival Location Y-axis Drivers",
     xlab = "Y-Location", breaks=20,
     ylab = "Frequency")
```

```{r}
nll <- function(mu, sigma) {
  
  -sum(log(dtruncnorm(drivers$y_coordinate, a = 0, b = 20, mean = mu, sd = sigma)))
}

fit_trunc <- mle(nll, start = list(mu = mean(drivers$y_coordinate), sigma = sd(drivers$y_coordinate)))
summary(fit_trunc)

mu_est <- fit_trunc@coef["mu"]       
sigma_est <- fit_trunc@coef["sigma"]


F_trunc <- function(x, mu, sigma) {
  lower <- 0
  upper <- 20
  (pnorm(x, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma)) /
    (pnorm(upper, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma))
}

ks_result <- ks.test(drivers$y_coordinate, function(x) F_trunc(x, mu_est, sigma_est))
print(ks_result)

# Plot histogram with density overlay in English
hist(drivers$y_coordinate, 
     col = "lightblue", border = "black",
     main = "Initial Driver Distribution on Y-axis", 
     xlab = "Y-Coordinate", 
     breaks = 20,
     ylab = "Density", 
     probability = TRUE)

# Plot the fitted truncated normal density curve
curve(dtruncnorm(x, a = 0, b = 20, mean = mu_est, sd = sigma_est), 
      from = 0, to = 20, 
      add = TRUE, col = "red", lwd = 2)

# Add legend with a blue square for observed data and red line with parameter estimates
legend("topleft", 
       legend = c("Observed Driver Location ", sprintf("Truncated Normal (mu = %.2f, sigma = %.2f)", mu_est, sigma_est)),
       col = c("lightblue", "red"),
       pch = c(15, NA),  
       lty = c(NA, 1),    
       lwd = c(NA, 2),
       cex=0.6)



```

```{r}


```

RIDERS

```{r}
#load datatset
riders <- read_excel("riders.xlsx", sheet = "Sheet1")

print(head(riders))
print(nrow(riders))
```

Extract the variables:

```{r}

#Inter-arrival time for driver (nrow-1 observation!)
riders$inter_arr <- c(NA,diff(riders$request_datetime)/3600) #in hour

riders$ride_time <- as.numeric(difftime(riders$dropoff_datetime,
                                           riders$pickup_datetime, units="hours"))

#Arrival coordinates:

#x-axis PICK UP
riders$x_coordinate_pu <- as.numeric(gsub("[()]", "", 
                            sapply(strsplit(riders$pickup_location, ", "), "[", 1)))

#y-axis PICK UP
riders$y_coordinate_pu <- as.numeric(gsub("[()]", "", 
                            sapply(strsplit(riders$pickup_location, ", "), "[", 2)))

#x-axis DROP OFF
riders$x_coordinate_do <- as.numeric(gsub("[()]", "", 
                            sapply(strsplit(riders$dropoff_location, ", "), "[", 1)))

#y-axis DROP OFF
riders$y_coordinate_do <- as.numeric(gsub("[()]", "", 
                            sapply(strsplit(riders$dropoff_location, ", "), "[", 2)))


#verify
head(riders[, c("id" ,"x_coordinate_pu", "y_coordinate_pu","x_coordinate_do", "y_coordinate_do", "inter_arr","ride_time")])
```

```{r}
#inter-arrival riders

#histogram for inter-arrival
hist(riders$inter_arr, col = "lightblue", border = "black", probability = TRUE,
     breaks=20, main = "Distribution of Inter Arrival Times for Riders", 
     xlab = "Inter-arrival time (h)",mylab = "Density")

#plot over the hist the candidate distribution
x_vals <- seq(0, max(na.omit(riders$inter_arr)), length.out = 100)
lines(x_vals, dexp(x_vals, rate = 30), col = "darkblue", lwd = 2)

legend("topright", legend = c("Observed Inter arrival times", 
                              "Exponential Distribution (rate 30)"),
       col = c("lightblue", "darkblue"), lty = c(NA, 1), pch = c(15, NA),
       lwd = c(NA, 2), bg = "white")
```

TEST INTER-ARRIVAL TIME FOR RIDERS

```{r}
n_riders <- nrow(riders)
#3040 observations, k = 300 seems reasonable

#Again inter-arrival time of riders can be assumed as independent, so we perform a chi-squared test

lambda <- 30
k <- 300 #as before

p <- 1/k

expected_i <- n_riders * p #greater than 10 (thumb rule is ok)

#build intervals with constant prob inside

num_int <- 0:(k-1)

bins<- c(-log( 1 - num_int / k) / lambda, Inf) #build k intervals with const
                                                #prob

#count O_i

counting <- cut(na.omit(riders$inter_arr), breaks = bins, 
                include.lowest = TRUE)

O_i <- table(counting)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k-1, lower.tail = TRUE)

chi_stat > chi_quant

print(chi_stat)
print(chi_quant)
#ACCEPT EXP (rate = 30 ) for riders inter arrival time

```

```{r}
#PICK-UP LOCATION
plot(riders$x_coordinate_pu, riders$y_coordinate_pu,
     col = rgb(0, 0, 1, 0.5), pch = 16,
     xlab = "X", ylab = "Y",
     main = "Pick-up Position Distribution of Riders")
grid()

```

```{r}
k <- 20
a <- 0
b <- 20

#bins on x and y
x_seq <- seq(a, b , length.out = k + 1)
y_seq <- seq(a, b , length.out = k + 1)


#count on x and y
x_count <- cut(riders$x_coordinate_pu, breaks = x_seq, include.lowest = TRUE)
y_count <- cut(riders$y_coordinate_pu, breaks = y_seq, include.lowest = TRUE)

obs_table <- table(x_count, y_count)

expected_i <- sum(obs_table) / (k * k)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k*k-1, lower.tail = TRUE)

chi_stat > chi_quant

print(chi_stat)

print(chi_quant) 
#CLEARLY NOT UNIFORM!
```

Does not really look uniform, let's analyse the separate variables for pick up:

```{r}
#X-axis pickup riders
hist(riders$x_coordinate_pu, col = "lightblue", border = "black",
     main = "Pick-up Location Distribution on X-axis for Riders", 
     xlab = "X-Location", breaks=20,
     ylab = "Frequency")
```

Not uniform at all.

TEST X-AXIS PICK UP LOCATION

```{r}
# x pu

library(truncnorm)
library(stats4)

nll <- function(mu, sigma) {
  
  -sum(log(dtruncnorm(riders$x_coordinate_pu, a = 0, b = 20, mean = mu, sd = sigma)))
}

fit_trunc <- mle(nll, start = list(mu = mean(riders$x_coordinate_pu), sigma = sd(riders$x_coordinate_pu)))
summary(fit_trunc)

mu_est <- fit_trunc@coef["mu"]       
sigma_est <- fit_trunc@coef["sigma"]


F_trunc <- function(x, mu, sigma) {
  lower <- 0
  upper <- 20
  (pnorm(x, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma)) /
    (pnorm(upper, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma))
}

ks_result <- ks.test(riders$x_coordinate_pu, function(x) F_trunc(x, mu_est, sigma_est))
print(ks_result)

# Plot histogram with density overlay in English
hist(riders$x_coordinate_pu, 
     col = "lightblue", border = "black",
     main = "Pick-up Location Distribution on X-axis", 
     xlab = "X-Coordinate", 
     breaks = 20,
     ylab = "Density", 
     probability = TRUE)

# Plot the fitted truncated normal density curve
curve(dtruncnorm(x, a = 0, b = 20, mean = mu_est, sd = sigma_est), 
      from = 0, to = 20, 
      add = TRUE, col = "red", lwd = 2)

# Add legend with a blue square for observed data and red line with parameter estimates
legend("topright", 
       legend = c("Observed Rider Location", sprintf("Truncated Normal (mu = %.2f, sigma = %.2f)", mu_est, sigma_est)),
       col = c("lightblue", "red"),
       pch = c(15, NA),  
       lty = c(NA, 1),    
       lwd = c(NA, 2),
       cex=0.8)
```

```{r}

```

```{r}
#Y-axis pickup riders
hist(riders$y_coordinate_pu, col = "lightblue", border = "black",
     main = "Pick-up Location Distribution on Y-axis for Riders",
     xlab = "Y-Location", breaks=20,
     ylab = "Frequency")
```

Not uniform.

```{r}
# y pu

nll <- function(mu, sigma) {
  
  -sum(log(dtruncnorm(riders$y_coordinate_pu, a = 0, b = 20, mean = mu, sd = sigma)))
}

fit_trunc <- mle(nll, start = list(mu = mean(riders$y_coordinate_pu), sigma = sd(riders$y_coordinate_pu)))
summary(fit_trunc)

mu_est <- fit_trunc@coef["mu"]       
sigma_est <- fit_trunc@coef["sigma"]


F_trunc <- function(x, mu, sigma) {
  lower <- 0
  upper <- 20
  (pnorm(x, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma)) /
    (pnorm(upper, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma))
}

ks_result <- ks.test(riders$y_coordinate_pu, function(x) F_trunc(x, mu_est, sigma_est))
print(ks_result)

# Plot histogram with density overlay in English
hist(riders$y_coordinate_pu, 
     col = "lightblue", border = "black",
     main = "Pick-up Location Distribution on Y-axis", 
     xlab = "Y-Coordinate", 
     breaks = 20,
     ylab = "Density", 
     probability = TRUE)

# Plot the fitted truncated normal density curve
curve(dtruncnorm(x, a = 0, b = 20, mean = mu_est, sd = sigma_est), 
      from = 0, to = 20, 
      add = TRUE, col = "red", lwd = 2)

# Add legend with a blue square for observed data and red line with parameter estimates
legend("topright", 
       legend = c("Observed Rider Location", sprintf("Truncated Normal (mu = %.2f, sigma = %.2f)", mu_est, sigma_est)),
       col = c("lightblue", "red"),
       pch = c(15, NA),  
       lty = c(NA, 1),    
       lwd = c(NA, 2),
       cex=0.7)
```

```         
```

```         
```

Drop-off:

```{r}
#DROP-OFF
plot(riders$x_coordinate_do, riders$y_coordinate_do,
     col = rgb(0, 0, 1, 0.5), pch = 16,
     xlab = "X", ylab = "Y",
     main = "Drop-Off Position Distribution of Riders")
grid()

```

does not look uniform (again normal?)

```{r}
k <- 20
a <- 0
b <- 20

#bins on x and y
x_seq <- seq(a, b , length.out = k + 1)
y_seq <- seq(a, b , length.out = k + 1)


#count on x and y
x_count <- cut(riders$x_coordinate_do, breaks = x_seq, include.lowest = TRUE)
y_count <- cut(riders$y_coordinate_do, breaks = y_seq, include.lowest = TRUE)

obs_table <- table(x_count, y_count)

expected_i <- sum(obs_table) / (k * k)

chi_stat <- sum((O_i - expected_i)^2/expected_i)

chi_quant <- qchisq(1 - alpha, k*k-1, lower.tail = TRUE)

chi_stat > chi_quant

print(chi_stat)

print(chi_quant)

#NOT UNIFORM
```

```{r}
hist(riders$x_coordinate_do, col = "lightblue", border = "black",
     main = "Drop-off Location Distribution on X-axis for Riders",
     xlab = "X-Location", breaks=20,
     ylab = "Frequency")
```

Normal?

```{r}
nll <- function(mu, sigma) {
  
  -sum(log(dtruncnorm(riders$x_coordinate_do, a = 0, b = 20, mean = mu, sd = sigma)))
}

fit_trunc <- mle(nll, start = list(mu = mean(riders$x_coordinate_do), sigma = sd(riders$x_coordinate_do)))
summary(fit_trunc)

mu_est <- fit_trunc@coef["mu"]       
sigma_est <- fit_trunc@coef["sigma"]


F_trunc <- function(x, mu, sigma) {
  lower <- 0
  upper <- 20
  (pnorm(x, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma)) /
    (pnorm(upper, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma))
}

ks_result <- ks.test(riders$x_coordinate_do, function(x) F_trunc(x, mu_est, sigma_est))
print(ks_result)

# Plot histogram with density overlay in English
hist(riders$x_coordinate_do, 
     col = "lightblue", border = "black",
     main = "Drop-off Location Distribution on X-axis", 
     xlab = "X-Coordinate", 
     breaks = 20,
     ylab = "Density", 
     probability = TRUE)

# Plot the fitted truncated normal density curve
curve(dtruncnorm(x, a = 0, b = 20, mean = mu_est, sd = sigma_est), 
      from = 0, to = 20, 
      add = TRUE, col = "red", lwd = 2)

# Add legend with a blue square for observed data and red line with parameter estimates
legend("topright", 
       legend = c("Observed Rider Location", sprintf("Truncated Normal (mu = %.2f, sigma = %.2f)", mu_est, sigma_est)),
       col = c("lightblue", "red"),
       pch = c(15, NA),  
       lty = c(NA, 1),    
       lwd = c(NA, 2),
       cex=0.6)
```

```{r}

```

```{r}
#Y-axis pickup riders
hist(riders$y_coordinate_do, col = "lightblue", border = "black",
     main = "Drop-Off Location Distribution on Y-axis for Riders", 
     xlab = "Y-Location", breaks=20,
     ylab = "Frequency")
```

Don't know what, but not uniform.

MAYBE WE COULD TRY AGAIN A NORMAL.

```{r}
#test again normality
nll <- function(mu, sigma) {
  
  -sum(log(dtruncnorm(riders$y_coordinate_do, a = 0, b = 20, mean = mu, sd = sigma)))
}

fit_trunc <- mle(nll, start = list(mu = mean(riders$y_coordinate_do), sigma = sd(riders$y_coordinate_do)))
summary(fit_trunc)

mu_est <- fit_trunc@coef["mu"]       
sigma_est <- fit_trunc@coef["sigma"]


F_trunc <- function(x, mu, sigma) {
  lower <- 0
  upper <- 20
  (pnorm(x, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma)) /
    (pnorm(upper, mean = mu, sd = sigma) - pnorm(lower, mean = mu, sd = sigma))
}

ks_result <- ks.test(riders$y_coordinate_do, function(x) F_trunc(x, mu_est, sigma_est))
print(ks_result)

# Plot histogram with density overlay in English
hist(riders$y_coordinate_do, 
     col = "lightblue", border = "black",
     main = "Drop-off Location Distribution on Y-axis", 
     xlab = "Y-Coordinate", 
     breaks = 20,
     ylab = "Density", 
     probability = TRUE)

# Plot the fitted truncated normal density curve
curve(dtruncnorm(x, a = 0, b = 20, mean = mu_est, sd = sigma_est), 
      from = 0, to = 20, 
      add = TRUE, col = "red", lwd = 2)

# Add legend with a blue square for observed data and red line with parameter estimates
legend("topleft", 
       legend = c("Observed Rider Location", sprintf("Truncated Normal (mu = %.2f, sigma = %.2f)", mu_est, sigma_est)),
       col = c("lightblue", "red"),
       pch = c(15, NA),  
       lty = c(NA, 1),    
       lwd = c(NA, 2),
       cex=0.6)

```

Now, we have a look to the time of the ride. (little bit more complicated since each time of the ride depends on the expected time mu_t).

Compute length of the ride (for riders that matched a driver and completed the ride, otherwise NA):

```{r}
# Default is NA
riders$ride_length <- NA

#Search for riders that started and finished the ride
valid_rides <- !is.na(riders$pickup_datetime) & !is.na(riders$dropoff_datetime)

#for that riders compute length od the ride
riders$ride_length[valid_rides] <- sqrt(
  (riders$x_coordinate_pu[valid_rides] - riders$x_coordinate_do[valid_rides])^2 + 
  (riders$y_coordinate_pu[valid_rides] - riders$y_coordinate_do[valid_rides])^2
)


# Verify
head(riders[, c("id", "x_coordinate_pu", "y_coordinate_pu", "x_coordinate_do", "y_coordinate_do", "ride_length")])

```

We have that mu_t=d_od/20, then time of the ride should be uniform between 0.8\*mu_t and 1.2\*mu_t:

```{r}
#mu_t is epected time of the ride
riders$expected_time <- riders$ride_length / 20

#lower and upper time for the ride
riders$lower_bound_time <- 0.8 * riders$expected_time
riders$upper_bound_time <- 1.2 * riders$expected_time

#Count how many of the rides have time that stay inside the bounds
within_bounds <- riders$ride_time[valid_rides] >= riders$lower_bound_time[valid_rides] &
                 riders$ride_time[valid_rides] <= riders$upper_bound_time[valid_rides]

#Percentange
percentage_within_bounds <- mean(within_bounds) * 100

cat("Percentage of the rides that have a time that is between 0.8 mu_t and 1.2 u_t is:",percentage_within_bounds,"%\n")
```

TEST:

```{r}
#normalise ride_time by mu_t in order to plot hist:
normal_ride_time <- riders$ride_time[valid_rides]/riders$expected_time[valid_rides]

#histogram
hist(normal_ride_time, col = "lightblue", border = "black",
     main = "Distribution of factor multiplying mu_t", xlab = "Factor",
     ylab = "Frequency", xlim=c(0.5,1.5))


#Add red line (paper says uniform between 5 and 8)
abline(v = 0.8, col = "red", lty = 2, lwd = 2)  
abline(v = 1.2, col = "red", lty = 2, lwd = 2)

legend("topright", legend = c("Observed Factors", "Uniform Limit (0.8,1.2)"),
       col = c("lightblue", "red"), lty = c(NA, 2), pch = c(15, NA), 
       lwd = c(NA, 2), bg = "white",cex=0.8)
```

In order to plot something "understandable" we normalise the real_time of the drive by the expected time. If the distribution is acutally uniform (0.8mu_t,1.2mu_t) then the normalised should be Uniform(0.2,1.2) which really seems to be the case.

We can then do a test on the normalised data! (to test H0: Uniform(0.8,1.2))

```{r}
#Since time of each ride can be dependent on others (example: peak hours all the rides can take longer than usual) this time we perform a Kolmogorov - Smirnov Test

a <- 0.8
b <- 1.2

n <- length(normal_ride_time)
F_n <- (1:n)/n 

F_true <- (sort(normal_ride_time) - a) / (b -a)

D <- max(abs(F_true - F_n)) #Correct!! same result as R function (see later)

C_0.9 <- 1.224 #FROM SLIDES

(sqrt(n) + 0.12 + 0.11/sqrt(n)) * D > C_0.9
```

```{r}

#double-check
ks_result <- ks.test(normal_ride_time, "punif", min = 0.8, max = 1.2)

print(ks_result)


```
